Using Visual Analytics to Explain\\[1ex] Black-Box Machine Learning

### 2018-04-05 tex/introduction/contribution.tex
user study findings

First, we generalized aggregating instance-level explanations to numerical input data.
We achieved this by using histograms of the input features, similar to \cite{seekaview}, ordered by the features' aggregated importance.
Furthermore, we allowed the comparison of selected, meaningful, subsets to each other.
Then, we compared this aggregated representation to the commonly used approach of individually inspecting instance-level explanations one-by-one using a tabular representation.

\todo{fill in latest work}
% In addition to those promising results, there is still work to do.
% It still needs to be shown that the proposed workflow of statistical analysis of instance-level explanations can be successfully applied to other data types than binary feature vectors, such as numerical features or highly redundant features such as found in images.
% Early results in this direction indicate that it might be beneficial to abandon the focus on features to explain machine learning models, but rather focus on instances directly.
% That is using instance-level explanations to obtain groups of \emph{instances} with similar behavior and then visualize those groups (\eg, using techniques proposed by \cite{seekaview}) in order to \emph{implicitly} explain them.
% This provides more flexible explanations than those proposed earlier as their capability goes beyond ``Feature A and feature B positively impact the prediction for this group of instances" to also allow insights of the form ``The number of circles formed by black pixels contribute positively to the prediction" which would be very difficult to formulate by a machine but are easily formulated by humans using visualizations.
% Furthermore, we are going to run a study on the implications on confidence in and trust of machine learning models when using different explanation approaches.
% \todo{END?}

>>>>

First, we generalized aggregating instance-level explanations to numerical input data.
We achieved this by using histograms of the input features, similar to \cite{seekaview}, ordered by the features' aggregated importance.
Furthermore, we allowed the comparison of selected, meaningful, subsets to each other.
Then, we compared this aggregated representation, in a user study, to the commonly used approach of individually inspecting instance-level explanations one-by-one using a tabular representation.

We found that aggregating instance-level explanations, with our method, significantly \textbf{outperforms} inspecting \textbf{individual explanations} in its ability to enable detection of biases in the data.
However, our method \textbf{requires explanations} in order to perform correctly.
Additionally, we found that instance-level explanations \textbf{hurt} bias detection performance for a \textbf{tabular representation}.
We hypothesize that the added difficulty of inspecting a table without the help of explanations forces users to \textbf{interact} more strongly with the user interface, thus providing better results.
This is a known effect (see Hullman~\etal~\cite{6064986}).

All in all, aggregating instance-level explanations perform as well as explanation-free tabular representations, which require high user engagement.
Additionally, our method promises a higher scalability due to its independence from the total number of instances.

Furthermore, we could reproduce findings from Stumpf~\etal~\cite{harmful}, claiming that instance-level explanations might make users over-confident in their understanding of a model.
However, we also showed that aggregated instance-level explanations overcome this problem.

% In addition to those promising results, there is still work to do.
% It still needs to be shown that the proposed workflow of statistical analysis of instance-level explanations can be successfully applied to other data types than binary feature vectors, such as numerical features or highly redundant features such as found in images.
% Early results in this direction indicate that it might be beneficial to abandon the focus on features to explain machine learning models, but rather focus on instances directly.
% That is using instance-level explanations to obtain groups of \emph{instances} with similar behavior and then visualize those groups (\eg, using techniques proposed by \cite{seekaview}) in order to \emph{implicitly} explain them.
% This provides more flexible explanations than those proposed earlier as their capability goes beyond ``Feature A and feature B positively impact the prediction for this group of instances" to also allow insights of the form ``The number of circles formed by black pixels contribute positively to the prediction" which would be very difficult to formulate by a machine but are easily formulated by humans using visualizations.
% Furthermore, we are going to run a study on the implications on confidence in and trust of machine learning models when using different explanation approaches.
% \todo{END?}
